---
title: "Workshop I 'Actuarial Science'"
author: "Amir_TALEBI"
date: "2023-12-10"
output: html_document
---

```{r}
library(tidyverse) 
library(dplyr)
library(ggplot2)

```

```{r}
insurance <- read.csv("/home/dvisual/insurance.csv", header = TRUE)
insurance
```

```{r}
summary(insurance[c('age', 'bmi', 'children', 'charges')])
```

```{r}
sapply(insurance, class)
```

```{r}
# to check for the outliers

for (i in 1:ncol(insurance)) {
  if (is.numeric(insurance[, i])) {
    boxplot(insurance[, i], outcol="Red", main=colnames(insurance)[i])
  }
}

```

```{r}
# to check for null values per column

colSums(is.na(insurance))
```
No null value. 

```{r}

sum(duplicated(insurance))

```

# so, there is one duplicated row in data frame. we remove it. 

```{r}

insurance <- unique(insurance)
insurance
```

```{r}

ggplot(insurance, aes(x = charges)) +
  geom_histogram(binwidth = 1500, fill = 'lightgreen', color = 'black') +
  labs(title='charges distribution', x = 'charges') +
  theme_minimal()

```
# long tail and as a result we have  outliers here.

```{r}

hist(insurance$age, main='Age distribution', xlab='Age', col='lightblue', border='black', breaks=15  , prob=TRUE)
lines(density(insurance$age), col = 'red')

```
# for the Age, that would be better if we had more data for the ages more than 50. Majority of records are from younger people. 

```{r}

insurance$sex <- as.factor(insurance$sex)

ggplot(insurance, aes(x=sex, fill=sex, group=sex)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(values=c('cyan', 'brown'))

```

```{r}

hist(insurance$bmi, main='BMI distribution', xlab='BMI', col='lightgreen', border='black', breaks=15 , prob=TRUE)
lines(density(insurance$bmi), col = 'red')

```
# BMI plot shows a gaussian distribution.

```{r}

ggplot(insurance, aes(x=children)) +
  geom_bar(fill='lightblue', color='black') +
  labs(title='Children counter', x='children') +
  theme_minimal()

```
# Children distribution looks interesting. With 0 and 5 as minority and majority of children in a family in our dataset.


```{r}

insurance$region <- as.factor(insurance$region)

ggplot(insurance, aes(x=region, fill=region)) +
  geom_bar(alpha=2.7) +
  labs(title='Region distribution') +
  theme_minimal()


```


```{r}
# group data by sex.

grouped_insurance_df <- insurance %>%
  group_by(sex)

summary_statistics <- grouped_insurance_df %>%
  summarise(
    min_charges = min(charges),
    max_charges = max(charges),
    mean_charges = mean(charges),
    count_charges = n(),
    std_charges = sd(charges)
  )

print(summary_statistics)
```
